### 2014-09-06 overview.tex
submitted version

\begin{table}
	\begin{center}
		\begin{tabular}{rcl}
			\multicolumn{1}{l}{{\tt s} $::=$} & & \\
			{\tt y = x.f} & $|$ & {\bf (heapr)} \\ 
			{\tt x.f = y}  & $|$ & {\bf (heapw)} \\ %\ $|$\ {\tt x.f = $c$}\
			{\tt z = x $bop$ y}\  & $|$ & {\bf (assign)} \\ %$|$\ {\tt z = $c$} $|$ {\tt z = new()}
			{\tt if (x $bop$  y) goto ...} & $|$ &  {\bf (branch)} \\
			{\tt lock(l)}\ $|$\ {\tt unlock(l)}  & $|$& {\bf (sync)} \\
			{\tt fork(t)}\ $|$\ {\tt join(t)}$|$ {\tt begin(t)}\ $|$\ {\tt end(t)} &  & {\bf (thread)}
		\end{tabular}
	\end{center}
	\caption{\label{Ta:syntax}Instruction Types}
\vspace{-0.15in}
\end{table}

@@

\item {\bf Intra-thread Value Constraints\ } The events should respect the 
constraints imposed by each instruction.
 For our running example, we obtain:
$$
\begin{array}{l}
	\theta'(W_{\tt x}^0) = 0 \wedge \theta'(W_{\tt y}^0) = 0 \wedge \theta'(W_{\tt y}^1) = 3 \\ 
	\wedge	\theta'(W_{\tt x}^2) = 1 \wedge \theta'(W_{\tt y}^3) = 5
	\end{array}
$$
As an example, $\theta'(W_{\tt y}^3)= 5$ denotes that the predicted run still assigns the value 5 to the variable {\tt y}  at event $e_3$.
\end{itemize}

@@

{\bf Inter-thread Value Constraints\ } These constraints capture what writes the read events read from and under what scheduling condition the value flow occurs.  The resulting formula for our example is
$$
\begin{array}{lcl}
	& & (\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^0) \wedge O'(e_4) < O'(e_1) ) \\
& \bigvee &
	(\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^1) \wedge O'(e_1) < O'(e_4) < O'(e_3)) \\
& \bigvee &
	(\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^3) \wedge O'(e_3) < O'(e_4))
\end{array}
$$    
Notice, importantly, that the formula associates the value constraints 
with the scheduling order constraints. 
As an example, 
$\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^1) \wedge O'(e_1) < O'(e_4) < O'(e_3)$ 
means that, in the predicted run, the read $e_4$ reads from the write $e_1$, 
under the condition that $e_4$ happens after $e_1$ and no other  
writes (such as $e_3$) interleave them.

>>>>

\begin{table}
{ 
	\begin{center}
		\begin{tabular}{rcl}
			\multicolumn{1}{l}{{\tt s} $::=$} & & \\
			{\tt y = x.f} & $|$ & {\bf (heapr)} \\ 
			{\tt x.f = y}  & $|$ & {\bf (heapw)} \\ %\ $|$\ {\tt x.f = $c$}\
			{\tt z = x $bop$ y}\  & $|$ & {\bf (assign)} \\ %$|$\ {\tt z = $c$} $|$ {\tt z = new()}
			{\tt if (x $bop$  y) goto ...} & $|$ &  {\bf (branch)} \\
			{\tt lock(l)}\ $|$\ {\tt unlock(l)}  & $|$& {\bf (sync)} \\
			{\tt fork(t)}\ $|$\ {\tt join(t)}$|$ {\tt begin(t)}\ $|$\ {\tt end(t)} &  & {\bf (thread)}
		\end{tabular}
	\end{center}
	}
	\caption{\label{Ta:syntax}Instruction Types}
\vspace{-3em}
\end{table}

@@

\item {\bf Intra-thread Value Constraints\ } The events should respect the 
constraints imposed by each instruction.
 For our running example, we obtain:
$$
{\small 
\begin{array}{l}
	\theta'(W_{\tt x}^0) = 0 \wedge \theta'(W_{\tt y}^0) = 0 \wedge \theta'(W_{\tt y}^1) = 3 \\ 
	\wedge	\theta'(W_{\tt x}^2) = 1 \wedge \theta'(W_{\tt y}^3) = 5
	\end{array}
}
$$
As an example, $\theta'(W_{\tt y}^3)= 5$ denotes that the predicted run still assigns the value 5 to the variable {\tt y}  at event $e_3$.
\end{itemize}

@@

{\bf Inter-thread Value Constraints\ } These constraints capture what writes the read events read from and under what scheduling condition the value flow occurs.  The resulting formula for our example is
$$
{\small 
\begin{array}{lcl}
	& & (\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^0) \wedge O'(e_4) < O'(e_1) ) \\
& \bigvee &
	(\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^1) \wedge O'(e_1) < O'(e_4) < O'(e_3)) \\
& \bigvee &
	(\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^3) \wedge O'(e_3) < O'(e_4))
\end{array}
}
$$    
Notice, importantly, that the formula associates the value constraints 
with the scheduling order constraints. 
As an example, 
$\theta'(R_{\tt y}^4)=\theta'(W_{\tt y}^1) \wedge O'(e_1) < O'(e_4) < O'(e_3)$ 
means that, in the predicted run, the read $e_4$ reads from the write $e_1$, 
under the condition that $e_4$ happens after $e_1$ and no other  
writes (such as $e_3$) interleave them.

### 2014-09-06 related.tex
submitted version

Given a concurrent execution trace $t$, PTA derives new traces to witness 
the data races. PTA is founded on the notion of sound causality, as it 
considers feasible reorderings of the input trace that prove a candidate 
data race as such.

Wang et al.~\cite{chao} pioneere the symbolic predictive analysis, 
which our approach also belongs to. However, the approach lacks support 
of heap encoding, therefore, cannot be applied to real-world applications. 
The first PTA that applies to real-world applications is proposed by 
Smagardakis et al. \cite{yannis}. In their original work, both synchronization 
constraints and inter-thread dependencies are preserved, where inter-thread 
dependencies are respected by only allowing reorderings that leave the 
dependence structure exhibited by the original trace intact. This ensures 
that the values of shared memory locations remain the same, which secures 
the soundness argument.

Said et al. \cite{Said:2011} and Huang et al. \cite{pldi14} proposed
PTAs that are also sound. Said et al. perform symbolic analysis of the 
input trace, and then utilize an SMT solver to search for schedules that 
establish the presence of data races. Soundness is guaranteed by their 
ability to precisely encode the semantics of sequential consistency. 
Inspired by Said et al., Huang et al. present an improvement, whereby 
control-flow information is encoded into the constraint system so as 
to relax flow dependencies as long as control dependencies are 
preserved. ExceptionNULL \cite{Farzan:2012} is another example of 
sound PTA. Its goal is to detect null dereferences rather than data races.

@@

\subsection{Maximality Claim}
Serbanuta et al. \cite{maximal} and Huang et al.~\cite{pldi14} both 
claim maximality of the detection capability, but under different assumptions. 
Assuming branch events are not modeled in the trace, \cite{maximal} guarantees maximality. Given few 
information in the trace, it conservatively requires the shared read to 
obtain the same value in the predicted run to ensure soundness. However, if 
branch events are modeled, the technique does not guarantee maximality.  
Assuming the branch events are modeled but local assignments are not recorded, 
\cite{pldi14} guarantees maximality. 
Compared to \cite{maximal}, it allows some events to read different 
values if they do not affect the following branch decisions. However, 
it requires the events that affect the branch decisions to read exactly 
the same values. In practice, these assumptions are too restrictive  
because local assignments can be easily collected in practice.
%the testers often have full access to the source/bytecode and 
%can collect local assignments freely.

Assuming local assignments are available in the trace, both techniques 
cannot guarantee maximality. As illustrated in our example and discussed 
in Section~\ref{sec:guarantee}, our analysis allows higher detection capability
%, i.e., we allow the events to read different values even they flow into the branch events; we allow the exploration of un-executed paths, both of which are not allowed by the above techniques.

>>>>

Given a concurrent execution trace $\tau$, PTA derives new traces to witness 
the data races. PTA is founded on the notion of sound causality, as it 
considers feasible reorderings of the input trace that prove a candidate 
data race as such.

Wang et al.~\cite{chao} pioneer the symbolic predictive analysis, 
which our approach also belongs to. However, the approach lacks support 
of heap encoding, therefore, cannot be applied to real-world applications. 
The first PTA that applies to real-world applications is proposed by 
Smagardakis et al. \cite{yannis}. In their original work, both synchronization 
constraints and inter-thread dependencies are preserved, where inter-thread 
dependencies are respected by only allowing reorderings that leave the 
dependence structure exhibited by the original trace intact. This ensures 
that the values of shared memory locations remain the same, which secures 
the soundness argument.

Said et al. \cite{Said:2011} and Huang et al. \cite{pldi14} proposed
PTAs that are also sound. Said et al. perform symbolic analysis of the 
input trace, and then utilize an SMT solver to search for schedules that 
establish the presence of data races. Soundness is guaranteed by their 
ability to precisely encode the semantics of sequential consistency. 
Inspired by Said et al., Huang et al. present an improvement, whereby 
control-flow information is encoded into the constraint system so as 
to relax value flow dependencies as long as the values do not flow into the control-flow event.
 ExceptionNULL \cite{Farzan:2012} is another example of 
sound PTA. Its goal is to detect null dereferences rather than data races.

@@

\subsection{Maximality Claim}
Serbanuta et al. \cite{maximal} and Huang et al.~\cite{pldi14} both 
claim maximality of the detection capability, but under different assumptions. 
Assuming branch events are not modeled in the trace, \cite{maximal} guarantees maximality. Given few 
information in the trace, it conservatively requires the shared read to 
obtain the same value in the predicted run to ensure soundness. However, if 
branch events are modeled, the technique does not guarantee maximality.  
Assuming the branch events are modeled but local assignments are not recorded, 
\cite{pldi14} guarantees maximality. 
Compared to \cite{maximal}, it allows some events to read different 
values if the values do not flow into the following branch conditions, as they do not affect the following branch decisions. However, 
it requires the events that affect the branch decisions to read exactly 
the same values. In practice, these assumptions are too restrictive  
because local assignments can be easily collected in practice.
%the testers often have full access to the source/bytecode and 
%can collect local assignments freely.

Assuming local assignments are available in the trace, both techniques 
cannot guarantee maximality. As illustrated in our example and discussed 
in Section~\ref{sec:guarantee}, our analysis allows higher detection capability.
%, i.e., we allow the events to read different values even they flow into the branch events; we allow the exploration of un-executed paths, both of which are not allowed by the above techniques.
